---
title:           "Exploration of algorithms and parameters"
author:          Dennis Wollersheim
date:            2020-01-13
linkcolor:       cyan
citecolor:       grey
output:
  pdf_document:
    highlight:   zenburn
---

\tableofcontents

```{r initial}

source('lib/base_initialise.R')
source('lib/functions.R')
source('lib/gps_functions.R')
source('lib/location_prep.R')
source('lib/load_results.R')
source('lib/evaluate_staypoint_estimates_helper.R')
source('lib/survey_functions.R')
source("explore/failure_analysis_florian_surveys.R")
min_accuracy = 10
```


# Focus on single algorithm

Due to business rules, 20M radius is preferable, and due to the poor quality of the gold standard, low precision is less
of a worry than low recall

So, we focus on the best recall, in the 20M radius.

Optics distance algorithm, harshest accuracy filter on GPS points, 15 minute staypoint windows

```{r best_precision_and_recall}

df_results %>%
#  filter(  algorithm == 'optics.distance') %>%
#  filter(  filter_type == 'filtered.accuracy') %>%
#  filter( recall > .18 ) %>%
  #filter( max_sp_radius == '10'  ) %>%
  arrange( desc( recall )) %>%
  ggplot( aes( precision, recall) ) +
  geom_jitter(position = position_jitter(seed = 1)) +
  geom_text(aes( label=source ), position = position_jitter(seed = 1), check_overlap=TRUE)

df_results %>%
  arrange(desc(total_sp_duration)) %>%
  dplyr::select( source, predictable_hits, predictable_misses, sp_total, total_sp_duration)

df_results %>%
  filter(interpol_parm > 0 ) %>%
  ggplot( aes(sp_total, predictable_hits, color=interpol_parm, alpha=.1) ) +
  geom_jitter(position = position_jitter(seed = 1)) +
  geom_text(aes( label=source), color='black' , position = position_jitter(seed = 1), check_overlap=TRUE) +
  facet_grid( algorithm ~ max_sp_radius )



```

# Focus on single algorithm
## find analysis of locations that results with best recall with above parameters

df_best contains all the gps locations (possibly extrapolated) of all the discovered staypoints, for
optics_distance_14400_900_20_interpolated_locations_600_filtered_accuracy_10 dataset

```{r read_in_best_file}

df_results %>%
  #filter(  filter_type == 'filtered.accuracy') %>%
  #filter(  algorithm == 'optics.distance') %>%
  #filter( max_sp_radius == '10'  ) %>%
  arrange(desc(total_sp_duration)) %>%
  head(1) %>%
  pluck( 'base_file' ) %>%
  { . } -> df_best_base_filename

df_results %>%
  filter( base_file == df_best_base_filename ) %>%
  pluck('accuracy_filter') %>%
  as.integer() %>%
  { . } -> min_accuracy


df_best_base_filename %>%
  paste0( 'df_merged_staypoints_',.) %>%
  readd(character_only=TRUE) %>%
  { . } -> df_best

df_best_base_filename %>%
  paste0( 'df_matching_survey_categories_df_matching_survey_df_merged_staypoints_', .) %>%
  readd(character_only=TRUE) %>%
  { . } -> df_best_matches


df_best %>%
  group_by( userid, night, n_staypoint) %>%
  summarise( min_ts = min(timestamp), max_ts = max(timestamp)) %>%
  { . } -> df_best_sp


```

# failure analysis

What could go wrong with staypoints:

1) did not capture any gps at all for that night
1) poor qualtity GPS points
1) insufficient duration gps points
2) battery ran out earlier in the evening
1) survey points are all earlier than GPS points

## survey data analysis from pov of GPS data

```{r data_Prep}


readd( df_all_ts  ) %>%
  group_by( id, userid, night) %>%
  dplyr::summarise( timestamp = min(timestamp)) %>%
  { . } ->  df_surveys

df_surveys_to_match = get_surveys_to_match()

df_nights_with_gps = get_nights_with_gps()

```


## nights with surveys, but missing GPS entirely

```{r nights_without_gps}

df_surveys_to_match %>%
  anti_join( df_nights_with_gps) %>%
  count()


df_surveys_to_match %>%
  inner_join( df_nights_with_gps %>% dplyr::select(userid, night), by=c('userid','night')) %>%
  { . } -> df_surveys_to_match_2

```

#

# too small duration GPS
which survey nights had less than 20 minutes of GPS in a night

```{r too_little_gps}

df_nights_with_gps %>%
  filter( duration < 1200 ) %>%
  inner_join( df_surveys_to_match_2, by=c('userid','night')  ) %>%
  count()


df_surveys_to_match_2 %>%
  anti_join( df_nights_with_gps %>% filter( duration < 1200 ) ) %>%
  { . } -> df_surveys_to_match_3

```


## battery failure
which nights only had surveys AFTER the last GPS point (battery)

```{r no_battery}

df_surveys_to_match_3  %>%
  group_by( userid, night ) %>%
  dplyr::summarise( first_survey_timestamp = min( timestamp), last_survey_timestamp = max(timestamp)) %>%
  ungroup() %>%
  inner_join( df_nights_with_gps ) %>%
  filter( last_gps_timestamp < first_survey_timestamp )  %>%
  dplyr::select( userid, night ) %>%
  inner_join(df_surveys_to_match_3 )  %>%
  count()

df_surveys_to_match_3  %>%
  group_by( userid, night ) %>%
  dplyr::summarise( first_survey_timestamp = min( timestamp), last_survey_timestamp = max(timestamp)) %>%
  ungroup() %>%
  inner_join( df_nights_with_gps ) %>%
  filter( last_gps_timestamp < first_survey_timestamp )  %>%
  dplyr::select( userid, night ) %>%
  { . } -> df_lost_battery_nights

df_surveys_to_match_3  %>%
  anti_join( df_lost_battery_nights) %>%
  { . } -> df_surveys_to_match_4
```

# surveyes BEFORE GPS


 note, we have many more nights where the first survey happens PRIOR to to the first GPS point

 fully 1/2 of the surveys are prior to the first gps point

```{r surveys_prior}

df_surveys_to_match_4  %>%
  group_by( userid, night ) %>%
  dplyr::summarise( first_survey_timestamp = min( timestamp), last_survey_timestamp = max(timestamp)) %>%
  ungroup() %>%
  inner_join( df_nights_with_gps ) %>%
  filter( first_gps_timestamp > last_survey_timestamp )  %>%
  dplyr::select( userid, night ) %>%
  inner_join(df_surveys_to_match_4 ) %>%
  count()


df_surveys_to_match_4  %>%
  group_by( userid, night ) %>%
  dplyr::summarise( first_survey_timestamp = min( timestamp), last_survey_timestamp = max(timestamp)) %>%
  ungroup() %>%
  inner_join( df_nights_with_gps ) %>%
  filter( first_gps_timestamp > last_survey_timestamp )  %>%
  dplyr::select( userid, night ) %>%
{ . } -> df_surveys_prior_to_gps

df_surveys_to_match_4  %>%
  anti_join( df_surveys_prior_to_gps) %>%
  { . } -> df_surveys_to_match_5



```


# Notes from meeting 2020-01-20

predictable surveys
deserts - no surveys within
maximise correspondence between staypoints and predictable surveys
throwaway questionnaire before and after
throwaway questionnaire within the desert

must have 3 points
roc curve, maximise area under the curve
need to minimise staypoint area

correctly predicted predictable stationary picture  / # possibly predictable sp /  #staypoints


# find Deserts, eliminate staypoints in the deserts
335 desert eliiminated surveys, using a 10 minute window, and 5 minute leeway after last gps point

```{r find_deserts}


desert_length=10
desert_endpoint_offset = 5*60


readd(df_location) %>%
{ . } -> df_location

df_location %>%
  filter( accuracy < min_accuracy ) %>%
  group_by(userid, night) %>%
  dplyr::arrange(timestamp, .by_group=TRUE) %>%
  mutate( timestamp_start = lag(timestamp) + desert_endpoint_offset,
         timestamp_end = timestamp - desert_endpoint_offset,
         diff = (timestamp_end - timestamp_start)/60 ) %>%
  filter(!is.na(timestamp_start)) %>%
  filter( diff > desert_length )  %>%
  dplyr::select(userid, night, starts_with('timestamp_'), diff) %>%
  nest( deserts=c(starts_with('timestamp_'), diff)) %>%
  ungroup() %>%
  { . } -> df_location_deserts

#
# get_df_surveys_cleaned () %>%
#   { . } -> df_surveys_to_match_5

df_surveys_to_match_5 %>%
  group_by(userid, night) %>%
  mutate( timestamp_start = timestamp, timestamp_end = timestamp) %>%
  dplyr::select(userid, night, starts_with('timestamp_'), id) %>%
  nest( surveys=c(starts_with('timestamp_'), id)) %>%
  ungroup() %>%
  { . } -> df_surveys_to_match_nested


# keep track of ALL staypoints found so we don't lose any staypoints
# when we join them to the surveys in the next step
maximum_seconds_distant =  0

# which staypoints match survey timestamps
df_location_deserts %>%
  inner_join( df_surveys_to_match_nested, by=c('userid', 'night')) %>%
  group_by( userid, night ) %>%
  do( joined = interval_inner_join( .$surveys[[1]], .$deserts[[1]], by=c('timestamp_start','timestamp_end'),
                                   maxgap=maximum_seconds_distant ))  %>%
  unnest( joined ) %>%
  ungroup() %>%
  { . } -> df_surveys_in_desert

df_surveys_in_desert %>%  count()

# alternative way to count
df_surveys_to_match_nested%>%
  unnest(surveys) %>%
  full_join(  df_location_deserts %>% unnest(deserts), by=c('userid', 'night')) %>%
  filter(
        timestamp_start.x >= timestamp_start.y &
        timestamp_end.x <= timestamp_end.y ) %>%
  count()

df_surveys_to_match_5 %>%
  anti_join( df_surveys_in_desert, by='id') %>%
  { . } -> df_surveys_to_match_6

```

# get rid of surveys that are before or after the last gps points

```{r surveys_post}

df_surveys_to_match_6 %>%
  inner_join( df_nights_with_gps, by=c('userid', 'night')) %>%
  filter( timestamp < first_gps_timestamp - desert_endpoint_offset  | timestamp > last_gps_timestamp + desert_endpoint_offset ) %>%
  count()

df_surveys_to_match_6 %>%
  inner_join( df_nights_with_gps, by=c('userid', 'night')) %>%
  filter( timestamp < first_gps_timestamp - desert_endpoint_offset  | timestamp > last_gps_timestamp + desert_endpoint_offset ) %>%
  { . } -> df_surveys_pre_or_post_gps

df_surveys_to_match_6 %>%
  anti_join( df_surveys_pre_or_post_gps, by='id' ) %>%
  { . } -> df_surveys_to_match_7


df_surveys_to_match_final = df_surveys_to_match_7

get_predictable_surveys()

```

# failure analysis - wtf is happening with these lost surveys
# given the best location set, what surveys does it miss?

best location set gets about 25% of surveys

```{r failure}


df_surveys_to_match_final %>%
  distinct( userid, night)

df_best %>%
  distinct( userid, night, n_staypoint) %>% count()


df_best %>%
  distinct( userid, night) %>% count()

df_location %>%
  distinct( userid, night) %>% count()


df_surveys_to_match_final %>%
  anti_join( df_best_matches,
            by='id') %>%
            { . } -> df_surveys_missed

df_surveys_to_match_final %>%
  inner_join( df_best_matches,
           by='id') %>%
           { . } -> df_surveys_gotten

df_surveys_missed %>% count()
df_surveys_gotten %>% count()

#readd(df_matching_survey_categories_summary_all) %>%
#  filter( str_detect(source, paste0( df_best_base_filename, '$'))) %>%
#  distinct(source) %>%
#  pluck('source')
```

# Current failure rate

```{r fiaure_rate}

(df_surveys_missed %>% count()) / (df_surveys_gotten %>% count() + df_surveys_missed %>% count())



```

#  what are the characteristics of the missing surveys, for a single egrious example
map a single user

```{r single_user}

df_surveys_missed %>%
  count( userid, night, sort=TRUE, name='n_missing_surveys' ) %>%
  inner_join(  df_nights_with_gps,  n_gps_points, by=c('userid', 'night')) %>%
  mutate(hours = duration/3600, points_per_minute = n_gps_points/(duration/60) ) %>%
  { . } -> df_missing_summary


df_missing_summary %>% head(10)

df_missing_summary %>% count(userid) %>% count()

i=2
for( i in rownames(df_missing_summary %>% head(10)) ) {
  map_one_user(df_missing_summary [i,], df_location, df_survey, df_best, min_accuracy)
  plot_one_person_survey(df_missing_summary[i,], df_location,df_best_sp, df_surveys_missed, min_accuracy )
}



df_missing_summary %>%
  head(1) %>%
  inner_join(df_best, by =c("userid", "night") ) %>%
  distinct( n_staypoint) %>%
  count()


```

#what is the range of missing surveys per usernight

```{r what_is_wrong_with_survey_match}

df_missing_summary %>%
  ggplot( aes( n_missing_surveys)  ) + geom_histogram()



```

#  what is the difference between the time of the survey and the time of the GPS points, within staypoints

```{r surveytime_vs_gps_point_time}

df_best_sp  %>%
  group_by(userid, night) %>%
  summarise( timestamp = min(min_ts) ) %>%
  ungroup() %>%
  inner_join( df_surveys_missed , by = c('userid', 'night')) %>%
  mutate( diff =  timestamp.x - timestamp.y)  %>%
  ggplot( aes( diff/3600)) + geom_histogram()

df_best  %>%
  distinct(userid, night, n_staypoint)



```

