---
title:           "Exploration of algorithms and parameters"
author:          Dennis Wollersheim
date:            2020-01-13
linkcolor:       cyan
citecolor:       grey
output:
  pdf_document:
    highlight:   zenburn
---

\tableofcontents

```{r initial}

source('lib/base_initialise.R')
source('lib/functions.R')
source('lib/gps_functions.R')
source('lib/location_prep.R')
source('lib/load_results.R')
source('lib/evaluate_staypoint_estimates_helper.R')
source('lib/survey_functions.R')
source("explore/failure_analysis_florian_surveys.R")
min_accuracy = 10

display_one_staypoint_set = function( df, .userid, .night ) {

  readd(df, character_only=TRUE ) %>%
    filter( userid==.userid & night==.night) %>%
    mutate(
            m_lat = ll2m( latitude, min(latitude), m_per_latitude),
            m_lon = ll2m( longitude, min(longitude), m_per_longitude)) %>%
    ggplot( aes( m_lat, m_lon, color=as.factor(n_staypoint))) +
    ggtitle(paste(str_subset(.userid, '.*_'), .night, df)) +
    geom_point()
}

```
# demonstrate the difference between the different accuracy levels (100/10), and the impact on the different algorithms
for a single user/night

```{r compare_accuracies}

display_one_staypoint_set (
  'optics_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_10',
  '60df2368-deb6-4f46-aebc-4ad7dd80f434',
  '2014-10-31')

display_one_staypoint_set(
 'staypoints_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_10',
  '60df2368-deb6-4f46-aebc-4ad7dd80f434',
  '2014-10-31')

display_one_staypoint_set(
 'optics_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_100',
  '60df2368-deb6-4f46-aebc-4ad7dd80f434',
  '2014-10-31')

display_one_staypoint_set(
 'staypoints_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_100',
  '60df2368-deb6-4f46-aebc-4ad7dd80f434', '2014-10-31')


```

# match optics and staypoint, comparing found surveys behaviour
Note, when using a loose accuracy level, sp algorithm is slightly better at finding surveys

using dataset :
  optics_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_100

```{r compare_surveys}


readd( df_matching_survey_per_staypoint_df_matching_survey_optics_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_100) %>%
  mutate(algo='optics') %>%
  bind_rows( readd(df_matching_survey_per_staypoint_df_matching_survey_staypoints_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_100) %>%
            mutate(algo='sp')) %>%
  count( userid, night, algo) %>%
  spread(algo, n, fill=0) %>%
  arrange( optics - sp) %>%
  head(5 )

display_one_staypoint_set(
                          'optics_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_100',
                          '7907f345-ef4b-412a-9340-b56ebb589cca','2014-09-19')

display_one_staypoint_set(
                          'staypoints_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_100',
                          '7907f345-ef4b-412a-9340-b56ebb589cca','2014-09-19')


display_one_staypoint_set(
 'optics_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_10',
  '7907f345-ef4b-412a-9340-b56ebb589cca','2014-09-19')

display_one_staypoint_set(
 'staypoints_distance_14400_300_100_interpolated_locations_120_filtered_accuracy_10',
  '7907f345-ef4b-412a-9340-b56ebb589cca','2014-09-19')


```
# background

## what do the surveys look like
Florian was able to identify survey points at the following frequencies.
sp stands for if this survey was likely to be a staypoint (the 'gold' standard to match)
The best staypoint discovery algorithm will maximise precision and recall of sp type locations

```{r what_gets_most_surveys}

n_ptype_short_total() %>%
  inner_join( ptype_short()) %>%
  group_by( ptype_short, category ) %>%
  summarise( total_for_category=sum(n_ptype_loc)) %>%
  kableExtra::kable()

```

In summary, there are this breakdown of target staypoints
```{r breakdown}

  n_ptype_short_total() %>%
  inner_join( ptype_short()) %>%
  group_by( category ) %>%
  summarise( total_for_category=sum(n_ptype_loc)) %>%
  kableExtra::kable()

```

### which algorithim and parameters maximise discovery of appropriate surveys
Here are the top 20 algorithms, for raw survey discovery.  Optics is comparable, but both have very loose parameters

```{r top_20 }

df_results %>%
  arrange( desc( sp )) %>%
  mutate( pct_coverage = mkpct( sp / 1939 )) %>%
  dplyr::select( base_file, sp,pct_coverage ) %>%
  head(20) %>%
  kableExtra::kable()

df_results %>%
  arrange( desc( sp )) %>%
  mutate( pct_coverage = mkpct( sp / 1939 )) %>%
  dplyr::select( base_file, sp,pct_coverage, algorithm ) %>%
  head(20) %>%
  ggplot( aes( base_file, sp, fill=algorithm ) ) +
  geom_col() +
  coord_flip()


```
# Is there a difference between optics and distance staypoint discovery sets?
look at the characteristics of the best of each;  not much difference, at the wide parameter level

```{r algo_diff}

df_results %>%
  arrange( desc( sp )) %>%
  filter( str_detect( base_file, '^stay')) %>%
  head(1) %>%
  pluck('base_file') %>%
  paste0('df_matching_survey_', .) %>%
  readd( character_only=TRUE ) %>%
  { . } -> df_sp

df_results %>%
  arrange( desc( sp )) %>%
  filter( str_detect( base_file, '^optics')) %>%
  head(1) %>%
  pluck('base_file') %>%
  paste0('df_matching_survey_', .) %>%
  readd( character_only=TRUE ) %>%
  { . } -> df_op

#df_sp %>% count( userid, night )
#df_op %>% count( userid, night )
```


Maybe 1% difference in night overlap

```{r some_overlap }
df_op %>%
  anti_join( df_sp, by=c('userid','night'))

df_sp %>%
  anti_join( df_op, by=c('userid','night'))

```
# difference between optics and distance staypoint discovery sets at finer parameter levels  (staypoint radius=20m)
At this level, optics gets 13% fewer sp, same staypoint duration, many fewer staypoints

```{r finer_grain}

df_results %>%
  arrange( desc( sp )) %>%
  filter( max_sp_radius==20) %>%
  filter( str_detect( base_file, '^stay')) %>%
  head(1)

df_results %>%
  arrange( desc( sp )) %>%
  filter( max_sp_radius==20) %>%
  filter( str_detect( base_file, '^optics')) %>%
  head(1)

```

### 20M radius staypoint results
which what do results look like when looking at algorithim and parameters maximise discovery of appropriate surveys

Here are the top 20 algorithms, for raw survey discovery.  Optics is comparable, but both have very loose parameters

```{r  200m}

df_results %>%
  filter( max_sp_radius==20) %>%
  arrange( desc( sp )) %>%
  mutate( pct_coverage = mkpct( sp / 1939 )) %>%
  dplyr::select( base_file, sp,pct_coverage ) %>%
  head(20) %>%
  kableExtra::kable()

df_results %>%
  filter( max_sp_radius==20) %>%
  arrange( desc( sp )) %>%
  mutate( pct_coverage = mkpct( sp / 1939 )) %>%
  dplyr::select( base_file, sp,pct_coverage, algorithm ) %>%
  head(20) %>%
  ggplot( aes( base_file, sp, fill=algorithm ) ) +
  geom_col() +
  coord_flip()


```
# model precision / recall
we are trying to identify correct survey marked staypoints

- True positive - surveys identified as staypoints
- false positive - surveys missed by staypoint discovery
- true negative  - points neither marked by surveys nor identified as staypoints
- false negative  points identified as staypoints, but not marked by surveys (sp)

our gold standard is actuallly bronze

False negative criticism:  an algorithmm identified staypoint could still be a valid  staypoint

it might be an unsurveyed staypoint, maybe they didn't have a drink, or record it (e.g. visiting on the streetcorner)

we will get relativly good (but low) values for precision and recall if we assume that the surveypoints do identify staypoints


recall is the number of true positives divided by the number of true positives plus the number of false negatives.

Precision is defined as the number of true positives divided by the number of true positives plus the number of false positives.

##  what does precision and recall look like
duration in seconds, distance from centriod in meters

Optics is better.  larger radius is better
```{r overall_pr}

df_results %>%
  ggplot( aes( precision, recall, shape=algorithm, color= max_sp_radius  ) ) +
  geom_point()


df_results %>%
  ggplot( aes( precision, recall, shape=algorithm, color= min_sp_duration  ) ) +
  geom_point()

```

##  effect of different filtering algorithms

Excluding optics algorithm, because it only uses accuracy filter

Accuracy is the best, throw away the others
```{r effect_of_filtering_algorithm}


df_results %>%
  filter( algorithm != 'optics.distance') %>%
  ggplot( aes( precision, recall, shape=min_sp_duration, color= max_sp_radius  ) ) +
  geom_point() + facet_wrap( ~filter_type )



```

## effect of interpolation
Interpolation appears to have little effect

```{r effect_of_interpolation}

df_results %>%
  filter(  filter_type == 'filtered.accuracy') %>%
  ggplot( aes( precision, recall, shape=algorithm, color= max_sp_radius  ) ) +
  geom_point() + facet_wrap( ~interpol_parm )

```
