---
title: "Staypoint estimate analysis post Drake"
output:
  workflowr::wflow_html
---


# Determine which staypoint parameters work the best

## staypoint parameters 
The staypoint determination algorithm uses 4 variables;
 - min_staypoint_time  - minimum time, in minutes, that must stay within max_staypoint_distance
 * max_jump_time - maximum time, in minutes, between readings

## Tests for good staypoint parameters

 - See if they match the survey data timestamps - we assume that survey data timestamps correspond to staypoints
 * see if they match drinking establishment locations
 * see if they match home locations - todo later 4 

```{r intro, results='hide', message=FALSE, warning=FALSE, cache=FALSE}
options(warn=-1)
library( knitr )
opts_chunk$set(cache=TRUE, autodep=TRUE, eval=TRUE)

library(tidyverse)
library(drake)
library(wrapr)
library(skimr)
library(treemap)



```

```{r load_data, eval=TRUE}


# load in the individual locations information


```


# Reports - Staypoints evaluated according to geography

```{r show_staypoints_geography}

loadd(df_all_sp_match_survey ) 

df_all_sp_match_survey %>%
  group_by( source ) %>%
  summarise( sp_total = n(), surveys_total= sum(!is.na( which))) %>% 
  mutate( rate = surveys_total / sp_total) %>%
  arrange( desc( rate )) %>% 
  separate( col=source, 
           into=c(NA, NA, qc(type, min_staypoint_time, max_jump_time, max_staypoint_distance, rest)), 
           sep='_', 
           convert=TRUE, 
           extra='merge',
           remove=FALSE)  %>% 
           { . } -> df_geography_matches 

df_geography_matches  %>%
  filter( max_staypoint_distance < 40 ) %>%
  filter( max_jump_time > 120 ) %>%
  ggplot(  aes( sp_total, surveys_total, color=type )) +
  geom_point() + 
  facet_grid( min_staypoint_time ~ max_staypoint_distance)

loadd(df_all_sp_match_survey_mode)

df_all_sp_match_survey_mode %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, 'interpolated_', 'interpolated.')) %>%
  mutate( source = str_replace( source, 'filtered_', 'filtered.')) %>%
  mutate( source = str_replace( source, 'filtered_accuracy', 'filtered.accuracy')) %>% 
  { . } -> df2


df_all_sp_match_survey %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, '_', '.')) %>%
  mutate( source = str_replace( source, 'interpolated_', 'interpolated.')) %>%
  mutate( source = str_replace( source, 'filtered_', 'filtered.')) %>%
  mutate( source = str_replace( source, 'filtered_accuracy', 'filtered.accuracy')) %>% 
  { . } -> df1

df1 %>%  
  as.Node( pathName = 'source', pathDelimiter='_') %>% 
  { . } -> tree


df2 %>%  
  as.Node( pathName = 'source', pathDelimiter='_') %>% 
  { . } -> tree2


df1 %>% 
  separate( col=source, 
            into=c( paste('sp', 10:17, sep='') ), 
            sep='_', 
            convert=TRUE, 
            extra='merge',
            remove=FALSE)  %>% 
distinct( sp14) %>%


  dplyr::rename( interpol=ms17, filter=ms15) %>%
            { . } -> cols1


df2 %>% 
  separate( col=source, 
           into=c( paste('ms', 10:19, sep='') ), 
            sep='_', 
            convert=TRUE, 
            extra='merge',
            remove=FALSE)  %>% 
  dplyr::rename( interpol=ms17, filter=ms19) %>%
  dplyr::select(-ms16, -ms18, -ms10) %>%
  { . } -> cols2



df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total, fill=as.character(max_jump_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total/sp_total, fill=as.character(max_jump_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total/sp_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + 
  facet_wrap( .~max_staypoint_distance)


df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total, fill=as.character(max_jump_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total/sp_total, fill=as.character(max_jump_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total/sp_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + 
  facet_grid( type~max_staypoint_distance)


```



