---
title: "Staypoint estimate analysis"
output:
  workflowr::wflow_html
---


# Determine which staypoint parameters work the best

## staypoint parameters 
The staypoint determination algorithm uses 4 variables;
 - min_staypoint_time  - minimum time, in minutes, that must stay within max_staypoint_distance
 * max_jump_time - maximum time, in minutes, between readings

## Tests for good staypoint parameters

 - See if they match the survey data timestamps - we assume that survey data timestamps correspond to staypoints
 * see if they match drinking establishment locations
 * see if they match home locations - todo later 4 

```{r intro, results='hide', message=FALSE, warning=FALSE, cache=FALSE}
options(warn=-1)
library( knitr )
opts_chunk$set(cache=TRUE, autodep=TRUE, eval=TRUE)

library(tidyverse)
library(drake)
library(wrapr)



```

```{r load_data, eval=TRUE}


# load in the individual locations information

loadd(df_all_sp_match_survey ) 

```


# Reports - Staypoints evaluated according to geography

```{r show_staypoints_geography}

df_all_sp_match_survey %>%
  group_by( filename) %>%
  summarise( sp_total = n(), surveys_total= sum(!is.na( which))) %>% 
  mutate( rate = surveys_total / sp_total) %>%
  arrange( desc( rate )) %>% 
  separate( col=filename, 
           into=c(NA, NA, qc(type, min_staypoint_time, max_jump_time, max_staypoint_distance, rest)), 
           sep='_', 
           convert=TRUE, 
           extra='merge',
           remove=FALSE)  %>% 
           { . } -> df_geography_matches 

df_geography_matches  %>%
  filter( max_staypoint_distance < 40 ) %>%
  filter( max_jump_time > 120 ) %>%
  ggplot(  aes( sp_total, surveys_total, color=type )) +
  geom_point() + 
  facet_grid( min_staypoint_time ~ max_staypoint_distance)



df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total, fill=as.character(max_jump_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total/sp_total, fill=as.character(max_jump_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total/sp_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + 
  facet_wrap( .~max_staypoint_distance)


df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total, fill=as.character(max_jump_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total/sp_total, fill=as.character(max_jump_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total/sp_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + 
  facet_grid( type~max_staypoint_distance)


```
