---
title: "Staypoint estimate analysis post Drake"
output:
  workflowr::wflow_html
---


# Determine which staypoint parameters work the best

## staypoint parameters 
The staypoint determination algorithm uses 4 variables;
 - min_staypoint_time  - minimum time, in minutes, that must stay within max_staypoint_distance
 * max_jump_time - maximum time, in minutes, between readings

## Tests for good staypoint parameters

 - See if they match the survey data timestamps - we assume that survey data timestamps correspond to staypoints
 * see if they match drinking establishment locations
 * see if they match home locations - todo later 

```{r intro, results='hide', message=FALSE, warning=FALSE, cache=FALSE}
options(warn=-1)
library( knitr )
opts_chunk$set(cache=TRUE, autodep=TRUE, eval=TRUE)

library(tidyverse)
library(drake)
library(wrapr)
library(skimr)
library(treemap)
library(data.tree)



```

```{r load_data, eval=TRUE}

source('lib/load_data_post_drake.R')
# df_results

```
# what are the results:

  - start with N surveys
  - elminate duplicates
  - eliminate surveys with no GPS locations
  * match surveys with various staypoint algorithms
  * decide whch staypoint algorithms provide the most parsimonious resultset
  * match geographic locations with staypoint algorithms


## all surveys
 what is the distribution of survey types overall?

```{r }
loadd(df_all_ts )
df_all_ts %>%
  count( which )

df_all_ts %>%
  count( )

```
## what is the distribution of allowable survey types

```{r }

df_all_ts %>%
  filter( which %in% c('dq','env','forg','video')) %>%
  count( which )

df_all_ts %>%
  filter( which %in% c('dq','env','forg','video')) %>%
  count( )

```
## Valid surveys: 
what is the distribution of allowable survey types that have GPS locations

```{r }

loadd( df_all_ts_valid ) 

df_all_ts_valid  %>%
  count( which )

df_all_ts_valid  %>%
  count( )

```

# what exact surveys are matched
## second, distribution of surveys found by a single algorithm 

```{r }

df_results%>%
  ggplot( aes( sp_total, surveys_total, color=survey_rate)) +
  geom_point()

df_results %>%
  ggplot( aes( sp_total, surveys_total, color=filter_type)) +
  geom_point()+
  geom_text( aes( label=interpol))

df_results %>%
  ggplot( aes( sp_total, surveys_total, color=algorithm)) +
  geom_point()+
  geom_text( aes( label=interpol))

df_results %>%
  count( filter_type)


readd(df_matching_survey_mode_meanshift_mode_300_10_10_1_6_interpolated_locations_120_filtered_accuracy_100) %>%
  filter( !is.na(which_survey)) %>%
  mutate( surveys = str_replace_all(which_survey, '[0-9. -]*','')) %>%
  separate( surveys, paste0('a_', 0:9)) %>% 
  select( userid, night, starts_with('a_' )) %>%
  tidyr::gather( junk, survey, -userid, -night ) %>%
  filter( !is.na( survey )) %>%
  count( survey) 

```

# TODO - do some basic error checking on survey matching algorithm
# Dataset description

## matching staypoint determination algorithms
  3 parameters

##  noise filtering
  1) accuracy
  2) geohash
  3) sigma v1 and v2

## data augmentation
  * filter accuracy, then interpolation

## modeshift 
 5 parameters- 
  * filter accuracy, then interpolation

#  plan

measure percentage capture of all surveys

find best noise filtering
compare filter accuracy and filtered augmented accuracy

compare everything at once

compare different  match parameter sets


# compare algorithms

```{r plan, eval=TRUE}


df_results  %>%
  ggplot( aes( algorithm, survey_rate)) +
  geom_boxplot()


df_results  %>%
  ggplot( aes( algorithm, sp_total)) +
  geom_boxplot()

df_results  %>%
  mutate( magnitude = cut( surveys_total, 6 )) %>%
  select( algorithm, survey_rate, magnitude) %>%
  ggplot( aes( algorithm, survey_rate)) +
  geom_boxplot() +
  facet_wrap( ~magnitude ) + 
  ggtitle( "compare algorithms faceted by number of points retrieved")


df_results  %>%
  filter( !is.na( interpol)) %>%
  ggplot( aes( interpol, survey_rate)) +
  geom_boxplot() +
  facet_wrap( ~algorithm ) + 
  ggtitle( "compare algorithms faceted by number of points retrieved")




```


# Reports - Staypoints evaluated according to geography

```{r show_staypoints_geography}


df_all_sp_match_survey %>%
  group_by( source ) %>%
  summarise( sp_total = n(), surveys_total= sum(!is.na( which))) %>% 
  mutate( rate = surveys_total / sp_total) %>%
  arrange( desc( rate )) %>% 
  separate( col=source, 
           into=c(NA, NA, qc(type, min_staypoint_time, max_jump_time, max_staypoint_distance, rest)), 
           sep='_', 
           convert=TRUE, 
           extra='merge',
           remove=FALSE)  %>% 
           { . } -> df_geography_matches 

df_geography_matches  %>%
  filter( max_staypoint_distance < 40 ) %>%
  filter( max_jump_time > 120 ) %>%
  ggplot(  aes( sp_total, surveys_total, color=type )) +
  geom_point() + 
  facet_grid( min_staypoint_time ~ max_staypoint_distance)


df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total, fill=as.character(max_jump_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total/sp_total, fill=as.character(max_jump_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total/sp_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + 
  facet_wrap( .~max_staypoint_distance)


df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total, fill=as.character(max_jump_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot()

df_geography_matches %>%
  ggplot( aes(x=(max_jump_time), y=surveys_total/sp_total, fill=as.character(max_jump_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(max_staypoint_distance), y=surveys_total/sp_total, fill=as.character( max_staypoint_distance) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + facet_wrap(~type)

df_geography_matches %>%
  ggplot( aes(x=(min_staypoint_time), y=surveys_total/sp_total, fill=as.character( min_staypoint_time) )) +
  geom_boxplot() + 
  facet_grid( type~max_staypoint_distance)


```



