---
title:           "Exploration of algorithms and parameters"
author:          Dennis Wollersheim
date:            2020-01-13
linkcolor:       cyan
citecolor:       grey
output:
  pdf_document:
    highlight:   zenburn
---

\tableofcontents

```{r initial}

suppressWarnings( source('lib/base_initialise.R') )
source('lib/functions.R')
source('lib/gps_functions.R')
source('lib/location_prep.R')
source('lib/load_results.R')
source('lib/evaluate_staypoint_estimates_helper.R')
source('lib/survey_functions.R')
source("explore/failure_analysis_florian_surveys.R")
min_accuracy = 10
```


# all algorithms

```{r all}

df_results %>%
  filter(interpol_parm > 0  & algorithm == 'optics.distance') %>%
  ggplot( aes(total_sp_duration, predictable_hits, color=algorithm), alpha=.1 ) +
  geom_jitter(position = position_jitter(seed = 1)) +
  geom_text(aes( label=source), color='black' , position = position_jitter(seed = 1), check_overlap=TRUE) +
  facet_grid( min_sp_duration ~ max_sp_radius )


```


# Focus on single algorithm

it so happens that the largest duration staypoints in total gives the very high accuracy on predictable staypoints (99%)

so we use that algorithm.

## algorithm

optics_distance_14400_300_100_interpolated_locations_300_filtered_accuracy_100

## what do recsults look like

```{r results}

df_results %>%
  filter(base_file=='optics_distance_14400_300_100_interpolated_locations_300_filtered_accuracy_100') %>%
  kableExtra::kable()

```

# geography_matches
## How far is point from matched geography point

 filter any matches greater than 100M accuracy

```{r distance_to_matching_geography}

readd(df_matching_geography_df_summarise_staypoints_df_merged_staypoints_optics_distance_14400_300_100_interpolated_locations_300_filtered_accuracy_100) %>%
  { . } -> df

df %>%
  ggplot( aes( dist)) + geom_histogram()

df %>%
  filter( dist<100 ) %>%
  ggplot( aes( dist)) + geom_histogram()

```


## what types of staypoint geographic matches are there?


```{r geography_matches}

df %>%
  filter( dist<100 ) %>%
  count( type) %>%
  kableExtra::kable()

```

##  what is the sequence of geographic matches

```{r }



df %>%
  filter( dist<100 ) %>%
  group_by( userid, night) %>%
  arrange( ts_min, .by_group=TRUE ) %>%
  summarise( types = paste( type, collapse=','), number_of_identifiable_staypoints=n()) %>%
  ungroup() %>%
  count( types, number_of_identifiable_staypoints, sort=TRUE) %>%
  kableExtra::kable()


```

## what is the sequence, when we discard discareable matches

```{r }

df %>%
  filter( dist<100 & type != 'discard') %>%
  group_by( userid, night) %>%
  arrange( ts_min, .by_group=TRUE ) %>%
  summarise( types = paste( type, collapse=','), number_of_identifiable_staypoints=n()) %>%
  ungroup() %>%
  count( types, number_of_identifiable_staypoints, sort=TRUE) %>%
  kableExtra::kable()

```

## write raw data

```{r}

df %>%
  write_csv( 'data/florian_best_dataset.csv')

```



