---
title: "Explore Staypoint base Data Evaluation and Preperation"
output:
  workflowr::wflow_html
---


```{r dataPrep, results='hide', message=FALSE, warning=FALSE, cache=FALSE}
options(warn=-1)

source('lib/functions.R')
source('lib/get_data.R')
source('lib/location_prep.R')
library(tidyverse)
library(lubridate)

opts_chunk$set(cache=TRUE, autodep=TRUE)


```


```{r load_data}


df_location = get_df_location()
df_best_location = get_df_best_location( df_location )
df_all = get_df_all()
df_all_ts = get_df_all_ts( df_all )


```

# problems with the survey data
  - timezone is all over the place, is it valid?
  - we can't check it
  - we  have to assume it is all valid (because we are going to use it only to substantiate the location data)

# problems with the gps location data
  - timezone, timestamp and localtime are not consistent with each other

# what order do they do their timestamping in (given the timezone they say they are in  )
Ten most popular survey orderings


```{r}

df_all %>% 
  select( userid, day, id ) %>%
  inner_join( df_all_ts, by='id') %>%
  arrange( userid, day, ts ) %>%
  group_by( userid, id ) %>%
  summarise( ordr = paste( which, collapse='-')) %>%
  ungroup() %>%
  count( ordr, sort=TRUE ) %>% 
  head(10)


```

# what is the range of purported survey timestamp and gps location durations
```{r}

df_all %>% 
  mutate( id = row_number()) %>%
  select( userid, day, id ) %>%
  inner_join( df_all_ts, by='id') %>%
  arrange( userid, day, ts ) %>%
  group_by( userid, id ) %>%
  summarise( d=as.numeric( max( ts ) - min( ts), units='mins')) %>%
  filter( d<3000 ) %>%
  ggplot( aes( d)) + geom_histogram() + xlab("minutes, limited to 3000 ") + ggtitle( 'Survey duration (max timestamp-min)')


df_best_location %>%
  group_by( userid, night ) %>% 
  summarise( d = as.numeric( max( local_time ) - min( local_time ), units='mins' ),
            max( local_time ) , min( local_time ), 
            n=n()) %>%
    filter( d<500 ) %>%
  ggplot( aes( d)) + geom_histogram() + xlab("minutes, limited to 500 ") + ggtitle( 'GPS duration (max timestamp-min)')


df_all %>% 
  select( userid, night, id ) %>%
  inner_join( df_all_ts, by='id') %>%
  group_by( userid, night ) %>%
  summarise( night_min = min( date( ts ))
            , max_ts = max( ts )
            , min_ts = min( ts )
            , ordr = paste( which, collapse='-')) %>%
  ungroup() %>%
  { . } -> df_all_nights

df_all %>% 
  select( userid, day, id ) %>%
  inner_join( df_all_ts, by='id') %>%
  arrange( userid, day, ts ) %>%
  group_by( userid, id ) %>%
  summarise( ordr = paste( which, collapse='-')) %>%
  ungroup() %>%
  count( ordr, sort=TRUE ) %>% 
  head(10)


```

# what is the range of 'nights', and how do they match between the datasets

first, GPS traces with NO surveys
```{r}

df_best_location %>%
  distinct( userid, night ) %>%  
  anti_join( df_all_nights) %>%
  arrange( userid, night) %>%
  nrow()

```


next, number of survey evenings with no GPS traces :-(
```{r}


df_all_nights %>%
  anti_join( df_best_location ) %>%
  arrange( userid, night) %>%
  nrow()

```
# Survey timezones
We can assume survey timezones are correct, despite them being very disparate.  For example, 

```{r  survey_timezones  }

df_all_ts %>%
  count( timezone ) %>%
  arrange( desc( n), timezone) %>%
  head(20)

```

Sometime surveys from the same night are 2 or more timezones!


```{r explore_location_timezones_variety}

df_all %>% 
  select( userid, night, id ) %>%
  inner_join( df_all_ts, by='id') %>%
  group_by( userid, night ) %>%
  distinct( timezone, .keep_all=TRUE ) %>%
  inner_join( df_all_nights) %>%
  mutate( n=n()) %>%
  filter( n > 1) %>%
  arrange( desc( n), ts) %>%
  select( -night_min, -max_ts, -min_ts, -userid, -timestamp) 

```


# location data
timezone, time_stamp , and localtime all consistent (if we take DST into account).  We can use time_stamp


```{r explore_location_timezones}

df_location %>%
  mutate( lt = local_time %>% as.character() %>% ymd_hms( tz = "UTC") %>% seconds() - timestamp- timezone) %>% 
  count( lt )


```

# Survey data Final decision
- we will use timestamp, despite messy timezone



